---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  eval = nzchar(Sys.getenv("MAXIMA_R_README_EVAL")) # see note near the bottom
)
```

# maxima

<!-- badges: start -->
<!-- badges: end -->

maxima embeds [Maxima](https://maxima.sourceforge.io/), a general-purpose
computer algebra system, directly into R via
[ECL](https://ecl.common-lisp.dev/) (Embeddable Common Lisp), linked in
as a library rather than run as a subprocess, so calls into Maxima are
ordinary in-process function calls with no text-protocol round trips.

It is deliberately **thin**: its job is data-format interchange between R
and Maxima, and one generic mechanism ([`mx_call()`]) for calling any
Maxima function by name. There is no curated `integrate()`/`solve()`-style
wrapper API with its own argument-checking or defaults. That's left for
downstream packages to build on top, tailored to what they actually need.

## Installation

```r
# install.packages("pak")
pak::pak("mitchelloharawild/maxima")
```

There's nothing to install beyond R itself and a C/C++ toolchain (Rtools
on Windows; Xcode Command Line Tools on macOS; `build-essential` on
Debian/Ubuntu): **no system ECL and no system Maxima are required**.
This package vendors both and compiles them from source the first time
you install it, along with ECL's own bundled GMP, Boehm-GC and libffi.
That first install therefore takes a while (compiling Maxima's full
source tree is the long step); see `tools/versions.sh` for the exact
pinned versions being built.

Windows support (`configure.win`) is new and has not yet been verified
on a real Windows machine; see `inst/NOTES.md` if it fails to build.

## Example

```{r example}
library(maxima)

x <- mx_symbol("x")
x^2 + 1

mx_call("diff", x^2, x)
mx_call("integrate", x^2, x)
mx_call("solve", x^2 == 4, x)

sin(x)
as.double(mx_call("sqrt", 2))
```

Ordinary R arithmetic on `mx_expr` objects traces out a Maxima expression
tree (the same technique used by
[Ryacas](https://github.com/r-cas/ryacas) and
[caracas](https://github.com/r-cas/caracas)); [`mx_call()`] reaches
anything else: every Maxima function, because it never hard-codes which
ones exist.

```{r, echo = FALSE, results = "asis"}
if (!nzchar(Sys.getenv("MAXIMA_R_README_EVAL"))) {
  cat(
    "_(The chunks above aren't evaluated when this README is rendered ",
    "without a working build of the vendored Maxima core on hand, e.g. ",
    "on CRAN's or GitHub's own README-rendering infrastructure. Set ",
    "`MAXIMA_R_README_EVAL=true` to evaluate them for real.)_\n",
    sep = ""
  )
}
```
